---
title: "Analysis of LM03 Lightmeter Data"
author: "Richard Schweitzer"
date: "2023-02-13"
output: github_document
---

Do we want to output every figure as a svg instead of png? If so, determine here:
```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = 'svg') # set output device to svg
```

First, load some necessary packages and the LM03 data, which was pre-compiled from around 1400 individual trials. 
```{r}
# packages we'll need
library(data.table)
library(ggplot2)
library(signal)
# LM03 data (compiled previously)
load("LM03_data.rda")
str(lm03_data)
length(unique(lm03_data$filename)) # number of trials we've recorded
```

Now check out what the data looks like, use a random subset of data for that:
```{r, fig.width=16, fig.height=10}
subsample_now <- sample(lm03_data$filename)[1:20] # twenty random trials
# plot raw data:
ggplot(data = lm03_data[is.element(filename, subsample_now)], aes(x = time, y = pd1, color = "pd1")) + 
  geom_line(size = 1.5, alpha = 0.7) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now)], aes(x = time, y = pd2, color = "pd2"), 
            size = 1.5, alpha = 0.7) + 
  theme_minimal(base_size = 15) + 
  facet_wrap(~filename) + 
  labs(x = "Time [ms]", y = "Photodiode response [a.u.]", color = "Photodiode ID")
```

Looks reasonable, but to reduce the amount of data we could easily reduce the temporal range a bit.
```{r, fig.width=16, fig.height=10}
# reduce size of data:
lm03_data <- lm03_data[time>=200 & time <=400] 

# plot raw data again:
ggplot(data = lm03_data[is.element(filename, subsample_now)], aes(x = time, y = pd1, color = "pd1")) + 
  geom_line(size = 1.5, alpha = 0.7) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now)], aes(x = time, y = pd2, color = "pd2"), 
            size = 1.5, alpha = 0.7) + 
  theme_minimal(base_size = 15) + 
  facet_wrap(~filename) + 
  labs(x = "Time [ms]", y = "Photodiode response [a.u.]", color = "Photodiode ID")

# zoom in on single trials:
ggplot(data = lm03_data[is.element(filename, subsample_now[1:2])], aes(x = time, y = pd1, color = "pd1")) + 
  geom_line(size = 0.5, alpha = 0.7) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now[1:2])], aes(x = time, y = pd2, color = "pd2"), 
            size = 0.5, alpha = 0.7) + 
  theme_minimal(base_size = 15) + 
  facet_wrap(~filename) + 
  labs(x = "Time [ms]", y = "Photodiode response [a.u.]", color = "Photodiode ID")+ 
  coord_cartesian(xlim = c(250, 350))
```

There is quite a bit of noise in there, which we can smooth, right? Its likely that this is not strictly noise but in fact regular luminance modulations related to the frequency of the LEDs used in the Propixx projector instead of a classic color wheel. Still, even at 4000 Hz sampling rate of the LM03 light meter, this frequency is likely heavily undersampled. 
To make it easier for us, we'll smooth with a Savitzky-Golay filter...
```{r, fig.width=16, fig.height=10}
# smooth with sgolayfilt
lm03_data[ , pd1s := sgolayfilt(x = pd1, p = 3, n = 9), by = .(filename)]
lm03_data[ , pd2s := sgolayfilt(x = pd2, p = 3, n = 9), by = .(filename)]
# restrict to positive values
lm03_data[pd1s<0, pd1s := 0]
lm03_data[pd2s<0, pd2s := 0]

# compare raw and filtered results in single trial basis
ggplot(data = lm03_data[is.element(filename, subsample_now[1:2])], aes(x = time, y = pd1s, color = "pd1")) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now[1:2])], aes(x = time, y = pd1, color = "pd1"), 
            size = 1, alpha = 0.5) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now[1:2])], aes(x = time, y = pd2, color = "pd2"), 
             size = 1, alpha = 0.5) + 
  geom_line(size = 1.5, alpha = 1) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now[1:2])], aes(x = time, y = pd2s, color = "pd2"), 
            size = 1.5, alpha = 1) + 
  theme_minimal(base_size = 15) + 
  facet_wrap(~filename) + 
  labs(x = "Time [ms]", y = "Photodiode response [a.u.]", color = "Photodiode ID") + 
  coord_cartesian(xlim = c(250, 350))

# the entire subset, smoothed:
ggplot(data = lm03_data[is.element(filename, subsample_now)], aes(x = time, y = pd1s, color = "pd1")) + 
  geom_line(size = 1.5, alpha = 0.7) + 
  geom_line(data = lm03_data[is.element(filename, subsample_now)], aes(x = time, y = pd2s, color = "pd2"), 
            size = 1.5, alpha = 0.7) + 
  theme_minimal(base_size = 15) + 
  facet_wrap(~filename) + 
  labs(x = "Time [ms]", y = "Photodiode response [a.u.]", color = "Photodiode ID")
```

Okay, now we'll try to extract the center of gravity of these peaks. 
Since the data is still noisy, a simple max operation wouldn't be a great solution, so let's try a average weighted by luminance measurements. To make sure we exclude the irrelevant data points, we'll threshold first. 
Let's aggregate:
```{r, fig.width=16, fig.height=10}
cutoff <- 10 # ignore all data points below this value
# aggregate for each photo sensor, then merge
lm03_agg_pd1 <- lm03_data[pd1s > cutoff, .(center_pd1 = weighted.mean(x = time, w = pd1s)), 
                          by = .(filename, staAmp, spdFac, iniPos, movVel, sacDur)]
lm03_agg_pd2 <- lm03_data[pd2s > cutoff, .(center_pd2 = weighted.mean(x = time, w = pd2s)), 
                          by = .(filename, staAmp, spdFac, iniPos, movVel, sacDur)]
lm03_agg <- merge.data.table(x = lm03_agg_pd1, y = lm03_agg_pd2)
rm(lm03_agg_pd1, lm03_agg_pd2) # not needed anymore
# this is the absolute time passed between the two estimated peaks
lm03_agg[ , time_between := abs(center_pd1 - center_pd2)] 
lm03_agg[ , direction := ordered(iniPos)]
head(lm03_agg)

# plot histogram real quick:
ggplot(lm03_agg, aes(x = time_between, color = direction, fill = direction)) + 
  geom_histogram(binwidth = 1, position = "stack") +
  facet_grid(staAmp~spdFac) + 
  theme_minimal(base_size = 15) + 
  scale_color_viridis_d(end = 0.7) + scale_fill_viridis_d(end = 0.7) + 
  labs(x = "Time between photodiode responses [ms]", y = "Number of trials", 
       color = "Motion direction", fill = "Motion direction") 
```

This looks very clean. Now we can make a final aggregate and plot. 
Importantly, the two sensors were approximately 2 degrees of visual angle apart from each other. Since the velocity was constant, we can also evaluate whether the measured durations are compatible with the specified velocities.

```{r, fig.width=16, fig.height=6}
# first, make the population aggregate
lm03_superagg <- lm03_agg[ , .(time_between = mean(time_between), 
                               time_between_sd = sd(time_between), 
                               N = length(time_between)), 
                           by = .(staAmp, spdFac, movVel, sacDur, direction)]
lm03_superagg[ , time_between_se := time_between_sd / sqrt(N)]
# compute the expected duration given the distance between the two sensors
lm03_superagg[ , movVelAbs := spdFac * movVel]
lm03_superagg[ , assumedDist := 2]
lm03_superagg[ , assumed_time_between := assumedDist / movVelAbs * 1000]

# order:
lm03_superagg <- lm03_superagg[order(staAmp, spdFac, direction)]
head(lm03_superagg)

# save, so that this data can be plotted by someone else
write.csv(x = lm03_superagg, file = "LM03_results.csv", row.names = FALSE)

# final plot, dotted line indicates theoretically assumed duration
ggplot(data = lm03_superagg, aes(x = spdFac, y = time_between, color = direction)) + 
  geom_ribbon(data = lm03_superagg, aes(x = spdFac, y = time_between, fill = direction, 
                                          ymax = time_between + 2*time_between_sd, 
                                          ymin = time_between - 2*time_between_sd), 
                size = 1.5, alpha = 0.3, 
              color = NA) + 
  geom_line(alpha = 0.7, size = 1.5) + 
  geom_line(data = lm03_superagg, aes(x = spdFac, y = assumed_time_between, color = direction), 
            linetype = "dotted", size = 1.5, color = "black") + 
  facet_wrap(~staAmp, nrow = 1) + 
  theme_minimal(base_size = 15) + 
  scale_color_viridis_d(end = 0.7) + scale_fill_viridis_d(end = 0.7) + 
  labs(x = "Velocity [prop. of peak velocity]", y = "Time between photodiode responses [ms]", 
       color = "Motion direction", fill = "Motion direction")
```

Finally, this shows that leftward and rightward motion take the same time when presented on the Propixx DLP projector at 1440 Hz. In addition, measured linear motion speed -- as indicated by the measured time between the activations of two horizontally aligned photo diodes -- is exactly as specified in the stimulus. 

